\documentclass[a4paper,fleqn,usenatbib]{mnras}  
\usepackage{newtxtext,newtxmath}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
%\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage[english]{babel}
\usepackage[dvipdfmx]{graphicx}
\usepackage{parskip}
%\usepackage[export]{adjustbox}
\pdfminorversion 4

%
\usepackage{subfig}
\usepackage{caption}
%\usepackage{subcaption}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
\usepackage{pseudocode}
%
% Enable mutiple title row on tables.
\usepackage{multirow,booktabs}
%%%%%%%%%%%
% 
% Dealing with ``too many unprocessed floats'' use it with \clearpage 
\usepackage[section] {placeins}

% Defining inentation
\newcommand{\defindent}{\setlength{\parindent}{4ex}}
\newcommand{\zeroindent}{\setlength{\parindent}{0ex}}

% For marginal notes
%\usepackage{marginnote}
%\usepackage{booktabs}

\title[Estimating the atmospheric parameters of M-type
  stars]{Estimates of the atmospheric parameters of M-type stars: a
  machine learning perspective.}

\author[L. M. Sarro et al.]
          {
L.M. Sarro$^{1}$\thanks{lsb@dia.uned.es},
J. Ordieres-Mer\'e$^{2}$,       
A. Bello-Garc\'ia$^{3}$, 
A. Gonz\'alez-Marcos$^{4}$, 
E. Solano$^{5}$
\\
$^{1}$Department of Artificial Intelligence,Universidad Nacional de Educaci\'{o}n a Distancia, c/ Juan del Rosal, 16, 28040 Madrid.\\
$^{2}$Universidad Polit\'{e}cnica de Madrid (UPM), ETSII. PMQ Research Group, Jos\'{e} Guti\'{e}rrez Abascal 2, 28006 Madrid, Spain. \\
$^{3}$Universidad de Oviedo, Construction and Manufacturing Engineering Department, Campus de Viesques s/n, Gij\'{o}n, Asturias, Spain. \\
$^{4}$Universidad de la Rioja, P2ML Research Group, San Jos\'{e} de Calasanz 31, 26004 Logro\~{n}o, La Rioja, Spain. \\
$^{5}$Centro de Astrobiolog\'{\i}a (CSIC-INTA), Ctra. Ajalvir km 4, E-28850 Torrejón de Ardoz, Madrid, Spain.\\
}
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2017}

\begin{document} 
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

\begin{abstract}
    Estimating atmospheric parameters of M type stars has been a
    difficult task due to the lack of simple diagnostics in the
    stellar spectra.  We aim at uncovering good sets of predictive
    features of the stellar atmospheric parameters ($T_{\rm eff}$,
    $\log(g)$, [$M/H$]) in spectra of M type stars. We define two types
    of potential features (equivalent widths and integrated flux
    ratios) able to explain the atmospheric physical parameters. We
    search the space of feature sets using a Genetic Algorithm that
    evaluates solutions by their prediction performance in the
    framework of the BT-Settl library of stellar spectra. Thereafter,
    we construct eight regression models using different machine
    learning techniques and compare their performances with those
    obtained using the classical $\chi^2$ approach and Independent
    Component Analysis (ICA) coefficients. Finally, we validate the
    various alternatives using two sets of real spectra from the IRTF
    and Dwarf Archive collections (DA).  We find that the cross-validation
    errors are poor measures of the performance of regression models
    in the context of physical parameter prediction in M type
    stars. For R$\sim$2000 spectra of the signal-to-noise ratios
    typical of the IRTF and DA, feature
    selection with Genetic Algorithms (GA) or alternative techniques
    produces only marginal advantages with respect to representation
    spaces that are unconstrained in wavelength (full spectrum or
    ICA). We make available the atmospheric parameters for the two
    collections of observed spectra sd online material.
    
\end{abstract}


\begin{keywords}
Methods: data analysis; Methods: statistical;
     Techniques: spectroscopic; Stars: atmospheres; Stars: fundamental
     parameters; Stars: late-type; Stars: statistics;
\end{keywords}

%

\defindent{}

\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Methodology.}
\label{sec:meth}
\input{meth}

\section{Physical parameters of the IRTF collection of spectra.}
\label{sec:irtf}

In the following, we will summarise the results obtained for the IRTF
data set. We deal with the different physical parameters in separate
Sections. We start by reporting the Root Mean/Median Square Errors
(RMSE/RMDSE) with respect to the parameters gathered from the
literature by \cite{cesetti} and included in their Table 3.

\subsection{Spectral bands selected}
\input{irtf.fs}
\subsection{Regression models}
\input{irtf.rm}

\section{Physical parameters of the DA collection of spectra.}
\label{sec:ipac}

\subsection{Spectral bands selected}
\input{ipac.fs}
\subsection{Regression models}
\input{ipac.rm}

\section{Summary and conclusions}
\label{sec:summary}
\input{summary}

\section*{Acknowledgements}
This research has benefitted from the M, L, T, and Y dwarf compendium
of spectra housed at DwarfArchives.org.  The authors would like to
acknowledge funding by the Spanish Ministry of Economy and Innovation
through the grant AyA2014-55216. The IRTF library is provided by the
University of Hawaii under Cooperative Agreement no. NNX-08AE38A with
the National Aeronautics and Space Administration, Science Mission
Directorate, Planetary Astronomy Program. This research has made
extensive use of the R software \cite{R} and the {\it caret} R
package. Finally, 
the authors acknowledge the computer resources and technical assistance 
provided by the Centro de Supercomputaci\'on y Visualizaci\'on de Madrid (CeSViMa).

%-------------------------------------------------------------------


\begin{appendix}
\label{app}
\input{appendix-wavelengths}
\input{appendix-RMSE}
\input{appendix-figures}
\end{appendix}


\bibliography{references_m}{}
\bibliographystyle{mnras}

\end{document}
