%                                                                 aa.dem
% AA vers. 8.2, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[rnote]{aa} % for the research notes
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
% according to the author-year natbib style

%
\documentclass[printer]{aa}  

%
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
\usepackage{pseudocode}
%
% Enable mutiple title row on tables.
\usepackage{multirow}
%%%%%%%%%%%
% 
% Dealing with ``too many unprocessed floats'' use it with \clearpage 
\usepackage[section] {placeins}

% For marginal notes
\usepackage{marginnote}
\usepackage{booktabs}

\begin{document} 


   \title{Physical parameter estimates of M-type stars: a machine learning perspective.}

   \author{
          L. M. Sarro \inst{1}  
          \and
          J. Ordieres-Mere \inst{2} 
          \and
          A. Bello-Garcia\inst{3} 
          \and
          A. Gonzalez-Marcos\inst{4} 
          \and
          M.B. Prendes-Gero\inst{3} 
          }

   \institute{ 
      \inst{1} Universidad Nacional de Educaci\'{o}n a Distancia, \\
               Department of Artificial Intelligence.
               \email{lsb@uned.es}
        \and 
      \inst{2}  Universidad Polit\'{e}cnica de Madrid (UPM), PMQ Research Group, \\
              Jos\'{e} Guti\'{e}rrez Abascal 2, 28006 Madrid, Spain. 
              \email{j.ordieres@upm.es}
        \and 
      \inst{3}   Universidad de Oviedo, Construction and Manufacturing Engineering Department, \\
              Campus de Viesques s/n, Gij\'{o}n, Asturias, Spain. 
              \email{\{abello,mbprendes\}@uniovi.es} 
        \and 
      \inst{4}  Universidad de la Rioja, P2ML Research Group, \\
              Luis de Ulloa 20, 26004 Logro\~{n}o, La Rioja, Spain. 
              \email{ana.gonzalez@unirioja.es}
      }

   \date{Received \today; accepted }

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory
 
   \abstract
%   context heading (optional)
    {Estimating physical parameters of M type stars has been a
      difficult task due to the lack of simple diagnostics in the
      stellar spectra.}
% aims heading (mandatory)
    {We aim at uncovering good sets of predictive features of the
      stellar atmospheric parameters ($T_{\rm eff}$, $\log(g)$, $M/H$)
      in spectra of M type stars. Each feature is defined by the
      wavelength ranges that contain the spectral information and the
      pseudo-continuum. A set of features is referred to as a
      representation space.}
%   % methods heading (mandatory)
    {We define two types of potential features (equivalent widths and
      integrated flux ratios) as predictors of the atmospheric
      physical parameters. We search the space of feature sets using a
      Genetic Algorithm that evaluates solutions by their prediction
      performance in the framework of the BT-Settl library of stellar
      spectra. Thereafter, we construct eight regression models using
      different machine learning techniques and compare their
      performances with two alternative representation spaces: the
      full spectrum as in the classical $\chi^2$ approach, and the
      Independent Component Analysis (ICA) coefficients. Finally, we
      validate the various alternatives using two sets of real spectra
      from the IRTF and Dwarf Archive collections.}
%   % results heading (mandatory)
    {We find that the cross-validation errors are poor measures of the
      performance of regression models in the context of physical
      parameter prediction in M type stars. For R~2000 spectra of the
      signal-to-noise ratios typical of the IRTF and Dwarf Archive
      collections, feature selection with Genetic Algorithms or
      alternative techniques produces only marginal advantages with
      respect to representation spaces that are unconstrained in
      wavelength (full spectrum or ICA). We provide physical
      parameters for the two collections of observed spectra.}
%   % conclusions heading (optional), leave it empty if necessary 
    {}

   \keywords{ Methods: data analysis; Methods: statistical;
     Techniques: spectroscopic; Stars: atmospheres; Stars: fundamental
     parameters; Stars: late-type; Stars: statistics;}

   \maketitle
%

\section{Introduction}
\label{sec:intro}
\input{intro}

% The importance of M stars
% The problem of estimation of stellar parameters in the M regime: bands, no continuum...
% who and what. Teff calibrations
% \cite{rajpurohit}
% \cite{amelia}

% Atlases 
%\cite{2013arXiv1306.3709B} % for lambda > 1.1 um :(
%\cite{2009ApJS..185..289R} %IRTF library of cool star

%\cite{2012ApJ...748...93R} % Metall. and Teff indicators in the K band! :(

% Summary of spectral diagnostics in the IR for M stars


\section{Methodology.}
\label{sec:meth}
\input{meth}

\section{Physical parameters of the IRTF collection of spectra.}
\label{sec:irtf}

In the following, we will summarise the results obtained for the IRTF
data set. We deal with the different physical paramenters in separate
Sections. We start by reporting the Root Mean/Median Square Errors
(RMSE/RMDSE) with respect to the parameters gathered from the
literature by \cite{cesetti} and included in their Table 3.

\subsection{Spectral bands selected}
\input{irtf.fs}
\subsection{Regression models}
\input{irtf.rm}

\section{Physical parameters of the IPAC collection of spectra.}
\label{sec:ipac}

\subsection{Spectral bands selected}
\input{ipac.fs}
\subsection{Regression models}
\input{ipac.rm}
\subsection{Comparison with previous feature sets}
\input{ipac.ces}

\section{Summary and conclusions}
\label{sec:summary}
\input{summary}

\begin{acknowledgements}
This research has benefitted from the M, L, T, and Y dwarf compendium
of spectra housed at DwarfArchives.org.  The authors would like to
acknowledge funding by the Spanish Ministry of Economy and Innovation
through the grant AyA2014-55216. The IRTF library is provided by the
University of Hawaii under Cooperative Agreement no. NNX-08AE38A with
the National Aeronautics and Space Administration, Science Mission
Directorate, Planetary Astronomy Program. This research has made
extensive use of the R software \cite{R} and the {\it caret} R
package.
\end{acknowledgements}

%-------------------------------------------------------------------

\bibliography{references_m}{}
\bibliographystyle{bibtex/aa}

\begin{appendix}
\label{app}
\input{appendix-wavelengths}
\input{appendix-RMSE}
\input{appendix-figures}
\end{appendix}

\end{document}
